---
title: "Ames_Housing_Data"
author: "Burton-Cordova & Thibeaux"
date: "2023-04-04"
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = "hide", eval = FALSE)
library(tidyverse)
library(ggplot2)
library(GGally)
library(caret)
```

Load Data
```{r Load Data, results='hide'}
#Load Test Data
test = read.csv('https://github.com/athibeaux/MSDS-DDS/raw/main/Project/test.csv', header = TRUE, fill = TRUE)

#Load Train Data
train = read.csv('https://github.com/athibeaux/MSDS-DDS/raw/main/Project/train.csv', header = TRUE, fill = NA)

#Summary of Train Data
summary(train)

#Select Relevant Columns and Neighborhoods for Analysis Question 1
C21 = train %>% select(GrLivArea,Neighborhood,SalePrice) %>% filter(Neighborhood == "NAmes" | Neighborhood == "Edwards" | Neighborhood == "BrkSide")

C21$Neighborhood <- as.factor(C21$Neighborhood)
summary(C21)

#Check and Remove NA's
sum(is.na(C21$GrLIvArea))
```

# Addressing Assumptions

## Linearity
```{r Linearity Assumption, results='hide'}
# Without Lines
C21 %>% ggplot(aes(GrLivArea, SalePrice, color = Neighborhood)) + geom_point() +
  xlab("Living Area in 100 Sq. Feet") + ylab("Sale Price") +
  ggtitle("Square Footage of Living Areas vs. Sales Price")

# With Lines for each Neighborhood
C21 %>% ggplot(aes(GrLivArea, SalePrice, color = Neighborhood)) + geom_point() +
  geom_smooth(method = "lm") + 
  xlab("Living Area in 100 Sq. Feet") + ylab("Sale Price") +
  ggtitle("Square Footage of Living Areas vs. Sales Price")
```

## Normality
```{r Normality Assumption, results='hide'}
ggpairs(C21)

# Histogram for Sale Price
C21 %>% ggplot() + geom_histogram(aes(x = SalePrice)) + 
  ggtitle("Distribution of Sale Price Variable") + xlab("Sale Price")

# Histogram for Living Area Square Footage
C21 %>% ggplot() + geom_histogram(aes(x = GrLivArea)) +
  ggtitle("Distribution of Living Area Square Footage") +
  xlab("Living Area in 100 sq. ft.")

# QQ Plot for Living Area Square Footage
C21 %>% ggplot() + geom_qq(aes(sample = GrLivArea)) + 
  ggtitle("QQ Plot for Living Area Square Footage") + ylab("Living Area Square Footage")

# QQ Plot for Sale Price
C21 %>% ggplot() + geom_qq(aes(sample = SalePrice)) + 
  ggtitle("QQ Plot for Sale Price") + ylab("Sale Price")

# Log Transformation on GrLivArea
C21$logLivArea = log(C21$GrLivArea)

# Histogram for Log-Transformed Living Area Square Footage
C21 %>% ggplot() + geom_histogram(aes(x = logLivArea)) +
  xlab("Living Area, log-transformed") +
  ggtitle("Distribution of log-transformed Living Area Square Footage")

# QQ Plot for Log-Transformed Living Area Square Footage
C21 %>% ggplot() + geom_qq(aes(sample = logLivArea)) + 
  ggtitle("QQ Plot for Living Area Square Footage, log transformed") + 
  ylab("Living Area, log-transformed square footage")

# Log Transformation on Sale Price
C21$logprice = log(C21$SalePrice)

# Histogram for Log-Transformed Sale Price
C21 %>% ggplot() + geom_histogram(aes(x = logprice)) +
  xlab("Sale Price, log-transformed")
  ggtitle("Distribution of log-transformed Sale Price")

# QQ Plot for Log-Transformed Living Area Square Footage
C21 %>% ggplot() + geom_qq(aes(sample = logprice)) + 
  ggtitle("QQ Plot for Sale Price, log transformed") + 
  ylab("Sale Price, log-transformed")


```

## Equal Variance
```{r Equal Variance Assumption, results='hide'}
# Non Transformed Data
rd <- lm(SalePrice ~ GrLivArea, data = C21)
g = rd$residuals
m<-mean(g)
std<-sqrt(var(g))
hist(g, density=20, breaks=20, prob=TRUE, col="red",
     xlab="Residuals",  
     main="Residual Histogram with Normal Distribution")
curve(dnorm(x, mean=m, sd=std), 
      col="darkblue", lwd=2, add=TRUE, yaxt="n")

# Display residual plots for non-transformed data
plot(rd)

# Notice cluster effect of residuals!

# Log-Log Model 

log_rd <- lm(logprice ~ logLivArea, data = C21)
log_g = log_rd$residuals
log_m<-mean(log_g)
log_std<-sqrt(var(log_g))
hist(log_g, density=20, breaks=20, prob=TRUE, col="red",
     xlab="Residuals",  
     main="Residual Histogram with Normal Distribution")
curve(dnorm(x, mean=log_m, sd=log_std), 
      col="darkblue", lwd=2, add=TRUE, yaxt="n")

# Display residual plots for log-transformed data
plot(log_rd)
```
We have a few high leverage residuals. Let's take a closer look.

```{r Outlier Examination, results='hide'}
# Examine the largest values in the explanatory variable
SortBySqFt <- C21[order(C21$GrLivArea, decreasing = TRUE),]
head(SortBySqFt)

# Wow! Two houses with >4000 square feet were sold for under $200k?? Either that's a mistake (like someone left out a zero) or I need the number of that real estate agent IMMEDIATELY so he can get me that kind of price on a mansion! If these two outliers are an valid observation, we may be looking at a short sale or foreclosure. Since these kinds of sales are not what Century 21 wants to measure, we feel confident in removing them from our dataset.

# Save complete dataset as a separate object in case we need it later
C21_full <- C21

# Find rows with outlier data
which(C21[ , 1] > 4000)

# Those are the same datapoints highlighted in our Residuals vs Leverage plot, so let's remove them.
C21 <- C21[-c(131,339),]
```
Now that we've removed the invalid data points, let's look again at the residuals.
```{r Residuals after Removing Invalid Data, results='hide'}
# Log-Log Model 

log_rd <- lm(logprice ~ logLivArea, data = C21)
log_g = log_rd$residuals
log_m<-mean(log_g)
log_std<-sqrt(var(log_g))
hist(log_g, density=20, breaks=20, prob=TRUE, col="red",
     xlab="Residuals",  
     main="Residual Histogram with Normal Distribution")
curve(dnorm(x, mean=log_m, sd=log_std), 
      col="darkblue", lwd=2, add=TRUE, yaxt="n")

# Display new residual plots
plot(log_rd)
```
Now let's take another look at the fits
```{r Refit Model, results='hide'}
# Parameter Estimates
# Overall
fit = lm(logprice ~ logLivArea, data = C21)
summary(fit)
confint(fit)

# Without Lines
C21 %>% ggplot(aes(logLivArea, logprice, color = Neighborhood)) +
  geom_point() +
  ylab("Sale Price, log transformed") + 
  xlab("Living Area Sq. Footage, log transformed") + 
  ggtitle("Square Footage of Living Areas vs. Sales Price")

# By Neighborhood
fit_hoods = C21 %>% group_by(Neighborhood) %>% do(model = lm(logprice ~ logLivArea, data = .))

# Brookside
summary(fit_hoods[[2]][[1]])
confint(fit_hoods[[2]][[1]])

# Edwards
summary(fit_hoods[[2]][[2]])
confint(fit_hoods[[2]][[2]])

# North Ames
summary(fit_hoods[[2]][[3]])
confint(fit_hoods[[2]][[3]])

# With Regression Line
C21 %>% ggplot(aes(logLivArea, logprice)) + 
  geom_point() + geom_smooth(method = "lm") +
  ylab("Sale Price, log transformed") + 
  xlab("Living Area Sq. Footage, log transformed") + 
  ggtitle("Square Footage of Living Areas vs. Sales Price")

# With Lines for each Neighborhood
C21 %>% ggplot(aes(logLivArea, logprice, color = Neighborhood)) + 
  geom_point() + geom_smooth(method = "lm") +
  ylab("Sale Price, log transformed") + 
  xlab("Living Area Sq. Footage, log transformed") + 
  ggtitle("Square Footage of Living Areas vs. Sales Price")
```
# Compare competing models
## Fit on non-transformed data
```{r Competing Models, results='hide'}
# Fit on non-transformed data
# Parameter Estimates
# Overall
fit = lm(SalePrice ~ GrLivArea, data = C21)
summary(fit)
confint(fit)

# Without Lines
C21 %>% ggplot(aes(GrLivArea, SalePrice, color = Neighborhood)) +
  geom_point() +
  ylab("Sale Price") + 
  xlab("Living Area Sq. Footage") + 
  ggtitle("Square Footage of Living Areas vs. Sales Price")

# By Neighborhood
fit_hoods = C21 %>% group_by(Neighborhood) %>% do(model = lm(SalePrice ~ GrLivArea, data = .))

# Brookside
summary(fit_hoods[[2]][[1]])
confint(fit_hoods[[2]][[1]])

# Edwards
summary(fit_hoods[[2]][[2]])
confint(fit_hoods[[2]][[2]])

# North Ames
summary(fit_hoods[[2]][[3]])
confint(fit_hoods[[2]][[3]])

# With Lines for each Neighborhood
C21 %>% ggplot(aes(GrLivArea, SalePrice, color = Neighborhood)) + 
  geom_point() + geom_smooth(method = "lm") +
  ylab("Sale Price") + 
  xlab("Living Area Sq. Footage") + 
  ggtitle("Square Footage of Living Areas vs. Sales Price")
```
## Hold Neighborhood Constant

```{r Model holding neighborhood constant, results='hide'}

# Transformed Data
fit = lm(logprice ~ logLivArea + Neighborhood, data = C21)
summary(fit)
confint(fit)

# Raw Data
fit = lm(SalePrice ~ GrLivArea + Neighborhood, data = C21)
summary(fit)
confint(fit)

```

## Internal CV as a measure for competing models

```{r Internal Cross Validation, results='hide'}
# LOOCV method from library(caret), transformed data

train(logprice ~ logLivArea, method = "lm", data = C21, trControl = trainControl(method = "LOOCV"))

# LOOCV method from library(caret), raw data

train(SalePrice ~ GrLivArea, method = "lm", data = C21, trControl = trainControl(method = "LOOCV"))

# LOOCV method from library(caret), transformed data, holding Neighborhood constant

train(logprice ~ logLivArea + Neighborhood, method = "lm", data = C21, trControl = trainControl(method = "LOOCV"))

# LOOCV method from library(caret), raw data, holding Neighborhood constant

train(SalePrice ~ GrLivArea + Neighborhood, method = "lm", data = C21, trControl = trainControl(method = "LOOCV"))
```

